Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-16 drafts/hybrid_zone.tex
edits

\documentclass[12pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{color}
\usepackage{amssymb}

@@

\title{Ancestry linkage in hybrid zones}
\author{}
\date{} % delete this line to display the current date

@@

%%% BEGIN DOCUMENT
\begin{document}

@@

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
\yb{Thus, \alisa{ancestries} originating from the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}. 
%In addition to steep clines in ancestry,
\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}  
\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} 
Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
\yb{The expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. 
\alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

>>>>

\documentclass[12pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{color}
\usepackage{amssymb}
\usepackage{setspace}

@@

\title{Blocks of extended ancestry around selected loci in hybrid zones}
\author{}
\date{} % delete this line to display the current date

@@

\usepackage{lineno}

\usepackage{authblk}

\begin{document}

\thispagestyle{empty}
\author[*]{Alisa Sedghifar}
\author[$\S$]{Yaniv Brandvain}
\author[$\dag$]{Peter Ralph}

\affil[*]{\small{Lewis-Sigler Institute for Integrative Genomics and Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, 08544}}
%\affil[2]{Department of Evolution and Ecology, University of California, Davis, Davis, California, United Sates of America}
\affil[$\S$]{\small{Department of Plant Biology, University of Minnesota, St.~Paul, Minnesota, 55108}}
\affil[$\dag$]{\small{Department of Molecular and Computational Biology, University of Southern California, Los Angeles, California, 90089}}

%%% BEGIN DOCUMENT

@@

\abstract{Hybrid zones formed between recently diverged populations offer an opportunity to better understand the mechanisms underlying reproductive isolation and the process of speciation. Many hybrid zones are characterized by some form of selection against hybrid genotypes, and here we use a combination of theory and simulations a to gain a better understanding of how this impacts patterns of introgression within the genome, and throughout space. Based on a simple model of single-locus under dominance, we find that ancestral blocks surrounding loci experiencing selection are longer than average, especially when removed from the center of the hybrid zone.  The findings presented here suggest a means for characterizing hybrid zones and identifying candidate targets of selection.}

\linenumbers
\doublespacing

\section*{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact \yb{before reproductive isolation is complete, gene flow is still possible. Thus, \alisa{ancestries} originating from each of the two diverged populations spread from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}. 
%In addition to steep clines in ancestry,
\yb{The recent exchange of genetic material between diverged populations leaves a distinct pattern of variation across the genome, in which long tracts of divergent haplotypes from each ancestral population  are broken up by historical recombination events.}

\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}. While these patterns of co-ancestry have been described in multi-locus systems, the  \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations have been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.  Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. Variation in selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton 1979}. The effect on patterns of ancestral block lengths is, however, less well explored theoretically and, \alisa{while empirical patterns of block length in hybrid zones have also not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.
